# Cleaning Data

This chapter describes the process I used to clean my Metadata sheet


## Loading initial data and packages

For this section, I used packages DBI and tidyverse.
```{r}
library(DBI)
library(tidyverse)
```

The following code was used to load my data files:
```{r}
Video_metadata<-read.csv("../../Homework/Data_files/Video_metadata.csv")

```
I changed two column names as they had periods, which could interfere with coding down the line:
```{r}
colnames(Video_metadata)[5]<-"Close_wide"
colnames(Video_metadata)[10]<-"Still_on_SD_card"

```
I also loaded a couple functions I wanted to use later on:
```{r}
rowShift <- function(x, shiftLen = 1L) {
  r <- (1L + shiftLen):(length(x) + shiftLen)
  r[r<1] <- NA
  return(x[r])
}
`%nin%` = Negate(`%in%`)
```
rowShift allows me to call a different row from a column when changing my "See previous row" into data that isn't dependent on order
%nin% is quicker than calling every possible integer and rather specifying the couple I do not want, similar to !=

## Cleaning the data

I first checked how many different names I used for my carcasses.
```{r}
unique(Video_metadata$Species)

```
I was able to use a long pipe to clean all my data. There are several chunks, and I have split it accordingly:
```{r}
Video_metadata <- Video_metadata %>% 
  mutate(Species = case_when(
    Species %in% c("Ewe", "Ewe ", "ewe") ~ "ewe",
    Species %in% c("Lamb", "lamb") ~ "lamb",
    Species %in% c("Neotoma", "chipmunk") ~ "rodent",
    Species %in% "raccoon" ~ "raccoon",
    Species %in% "carp" ~ "carp",
    Species %in% "calf" ~ "calf",
    Species %in% "cow" ~ "cow",
    Species %in% "badger" ~ "badger",
    Species %in% c("Deer", "deer") ~ "deer",
    Species %in% "cat" ~ "cat",
    Species %in% "" ~ "DOUBLE CHECK"
  )) # %>% 
```
This cleans up all my different "species" into recognizable carcass groups. I didn't want to just use sheep and cattle, but instead differentiating between adult/young since there is a large size difference (and although I have weight later on, this is helpful for recognition). I also included a "Double check" group since I have some blanks, that I haven't entered in yet.
```{r eval=FALSE}
mutate(Still_on_SD_card = case_when(
    Still_on_SD_card == "Yes" ~ "Yes",
    Still_on_SD_card == "Erased" ~ "Erased",
    SD_card == "" ~ "Probably erased"
  )) %>% 
```
The next chunk does a similar role, with the column identifying whether the video is still on an SD card. I reclassified all blanks as probably erased, as I have entered in most SD card numbers. However, I wanted to keep there separate from Erased since I'm not entirely certain about this and should go back to check.

```{r eval=FALSE}
mutate(Any_Video = case_when(
    Video %in% c("eh", "y") ~ "yes",
    Video == "n" ~ "no",
    Video == "see above" ~ rowShift(Video, -1)
  )) %>% 
relocate(Any_Video, .after = Video) %>% 
mutate(Any_Video = case_when(
    Video == "n" ~ "no",
    Video %in% c("eh", "y") ~ "yes"
  )) %>%
```

This is again similar. I had one column about whether I had video from the visit, but I was using this both for useful and any video (think grass swaying in the wind, with no visits from scavengers). I had called this "eh", and here I separate the eh into no for useful video, but yes for any video. I also relocated the Any_video next to Video so I could check my work and that everything looked correct.

```{r eval=FALSE}
mutate(Useful_Video = case_when(
    Video == "y" ~ "yes",
    Video %in% c("eh", "n") ~ "no"
  )) %>% 
```
After I transferred data from video to Useful_Video, I mutated the data again to conform to the previous "yes"/"no" rather than y/n that was in the Video column.

```{r eval=FALSE}
mutate(Close_wide = case_when(
    Visit %nin% c("NA","") & Close_wide == "wide" ~ "wide",
    Visit %nin% c("NA","") & Close_wide == "close" ~ "close",
    Visit %in% c("NA","") & Close_wide == "" ~ "NA",
    Visit %nin% c("NA","") & Close_wide %in% c("",".") ~ "only"
  ))
```

I used my "not in" function to find the Visits that had values (ie wasn't blank or NA). Close_wide describes camera placement, as I usually had a camera closer and one farther away (giving a wider angle view). 

If Visits is blank or NA that means I haven't gotten to it yet, and Close_wide should also be blank. However, if Close_wide is blank with a value for Visit, it means only one camera was there, so I should label it as "only". I briefly also used a period to designate "only" so I included that as well.

## Connecting to database
Now that the data is cleaned, I can connect it to the condors database and have the database run all the specified checks.
```{r eval=FALSE}
Condors <- dbConnect(RSQLite::SQLite(), "condors.db")
dbExecute(Condors, "CREATE TABLE metadata (
Video_ID INTEGER PRIMARY KEY AUTOINCREMENT,
Carcass char(3),
Description varchar(25),
Species varchar(15),
Visit varchar(5),
Close_wide varchar(6) CHECK (Close_wide IN ('only','close','wide', 'NA')),
Box_drive varchar(3) CHECK (Box_drive IN ('Yes','No','')),
External_Hard_Drive varchar(3) CHECK (External_Hard_drive IN ('Yes','No','')),
Scored_on_BORIS varchar(3) CHECK (Scored_on_BORIS IN ('Yes','No','')),
Need_to_time_offset char(3),
Still_on_SD_card varchar(3) CHECK (Still_on_SD_card IN ('Yes','Erased','Probably erased')),
SD_card varchar(10),
Video varchar(20),
Any_Video varchar(20),
Condors varchar(20),
Comments varchar(50),
Useful_Video varchar(20),
FOREIGN KEY(Carcass) REFERENCES Carcasses(ID_number)
);")
dbWriteTable(Condors,"metadata",Video_metadata,append=TRUE)

```
This ran well and all my data looks good!